API design for machine learning software: experiences from the scikit-learn project

### 2013-06-25 paper.tex
Smaller font size for bibliography.
Cosmit.
Add project URL and reset footnote counter.
Small improvements

@@

The \sklearn project \citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a general-
purpose high-level language. Among other things, it includes classical learning
algorithms, model selection tools and preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a classification or regression task, the goal is to find a mapping
between input variables $X_1, ..., X_p$, called features, and
some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
such samples.  As such, matrices of numerical values constitute a natural and
widely used representation in machine learning. In this view, each row in the
matrix then corresponds to a sample of the dataset and each column to one of the
variables of the problem.

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in the case of classification, or the mean squared
error in the case of regression. Not however that, from an API point of view,
the only guideline is that the \texttt{score} method returns a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also exposes a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it does exemplify API design for scalable processing of ``big data''
that takes a different approach than comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

\section{Future Work}
\label{sec:future_work}
There are several problems with the current scikit-learn interface which
will be resolved in the future.
At present, the library does not support some classical machine learning tasks,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Tasks like structured prediction or reinforcement learning are
considered off-scope for the project, since they would require quite different
data representations and APIs.

@@

\bibliographystyle{abbrvnat}
\DeclareRobustCommand{\VAN}[3]{#3}
\bibliography{paper}

>>>>

\setcounter{footnote}{0}

@@

The \sklearn project\footnote{\url{http://scikit-learn.org}}
\citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a general-
purpose high-level language. Among other things, it includes classical learning
algorithms; model evalutation, selection and evaluation tools; and
preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

In most machine learning tasks, data is modeled as a set of variables.  For
example, in a classification or regression task, the goal is to find a mapping
between input variables $X_1, ..., X_p$, called features, and
some output variable $Y$. A sample is then defined as a tuple of values $([x_1,
..., x_p]^\mathrm{T}, y)$ of these variables and a dataset as a collection of
such samples.  As such, matrices of numerical values constitute a natural and
a widely used representation in machine learning: each row in the
matrix corresponds to one sample of the dataset and each column to one
variable of the problem.

@@

Finally, predictors should also provide a \texttt{score} function to assess the
performance on some input data. In supervised learning estimators, this method
takes as inputs arrays \texttt{X\_test} and \texttt{y\_test} and typically
computes the coefficient of determination between \texttt{y\_test} and
\texttt{predict(X\_test)} in regression, or the accuracy
in classification. Not however that, from an API point of view,
the only guideline is that the \texttt{score} method returns a value
that quantifies the quality of its predictions (the higher, the better).
Unsupervised learning estimators may also exposes a \texttt{score} function
to compute, for instance, the likelihood of the given data under its model.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it exemplifies API design for scalable processing of ``big data''
that takes a different approach than comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

\section{Future work}
\label{sec:future_work}
There are several problems with the current scikit-learn interface which
will be resolved in the future.
At present, the library does not support some classical machine learning
algorithms,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Tasks like structured prediction or reinforcement learning are
considered off-scope for the project, since they would require quite different
data representations and APIs.

@@

{\small
\bibliographystyle{abbrvnat}
\DeclareRobustCommand{\VAN}[3]{#3}
\bibliography{paper}
}
